### 2024-06-05 thesis/input_files/analysis.tex
checked links urls

% kernel comparison
\subsection{Kernel function choice}
This section analyses which kernel function best fits to the characteristics of the load data.
The kernel functions considered are: Gaussian RBF, Absolute Laplacian, Matern 0.5, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process}{sklearn.gaussian\_process}.
The pool of data for this study has been created by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.
SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modelled by the European Centre for Medium-Range Weather Forecasts (ECMWF).
Therefore, in this experimental study, we used the whole data of $2020$ as the training set and then tested our kernels on the $2021$ data.
Note that there are two kinds of predictions for data from $2021$ onward, one for each of the emission scenarios RCP$4.5$ and RCP$8.5$. In this study, we considered the more conservative RCP$4.5$ scenario data.

>>>>

% kernel comparison
\subsection{Kernel function choice}
This section analyses which kernel function best fits to the characteristics of the load data.
The kernel functions considered are: Gaussian RBF, Absolute Laplacian, Matern 0.5, Matern 1.5, Matern 2.5, Linear, Periodic, Polynomial, Sigmoid, and Cosine. 
Details concerning the hyperparameter selection for the different kernels can be found in \href{https://scikit-learn.org/stable/api/sklearn.metrics.html}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/api/sklearn.gaussian_process.html}{sklearn.gaussian\_process.kernels}.
The pool of data for this study has been created by combining \href{https://zenodo.org/records/7907883}{SECURES-Met} data (predictors) \cite{Formayer2023} and the load data from \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts}.
SECURES-Met data consists of historical data up to the end of $2020$ while from 2021 onward, the data consists of forecasts modelled by the European Centre for Medium-Range Weather Forecasts (ECMWF).
Therefore, in this experimental study, we used the whole data of $2020$ as the training set and then tested our kernels on the $2021$ data.
Note that there are two kinds of predictions for data from $2021$ onward, one for each of the emission scenarios RCP$4.5$ and RCP$8.5$. In this study, we considered the more conservative RCP$4.5$ scenario data.

### 2024-06-05 thesis/input_files/implementation.tex
checked links urls

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented Python open source library, thus, the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialised by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point algorithm. This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus, we just have to provide the $Q$ matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a single matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with in built scikit-learn methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}. Functions supported by our kernel quantile regression include: Gaussian RBF, Absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels. 
Kernels from these two submodules use a different convention in the hyperparameters. The former has the lengthscale parameter in the numerator while the latter in the denominator. In order to meaningfully compare across different kernels we made them consistent in this aspect. For this purpose we adhere to the convention of having lengthscale hyperparameters in the denominator.

>>>>

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented Python open source library, thus, the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialised by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point algorithm. This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus, we just have to provide the $Q$ matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a single matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with in built scikit-learn methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/api/sklearn.metrics.html}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/api/sklearn.gaussian_process.html}{sklearn.gaussian\_process.kernels}. Functions supported by our kernel quantile regression include: Gaussian RBF, Absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels. 
Kernels from these two submodules use a different convention in the hyperparameters. The former has the lengthscale parameter in the numerator while the latter in the denominator. In order to meaningfully compare across different kernels we made them consistent in this aspect. For this purpose we adhere to the convention of having lengthscale hyperparameters in the denominator.
